Evaluation of +S2 sea ice forecast and some other better title

### 2024-12-13 analysis/paper/access-ice.qmd

---
title: "Untitled"
format: gfm
execute: 
  echo: false
  warning: false
  message: false
  cache: false
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
  - ../../_extensions/ute/search-replace/search-replace.lua
search-replace:
  +S2: ACCESS-S2
  +S1: ACCESS-S1
  +CDR: NOAA/NSIDC CDRV4
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)
library(tagger)

cdo_options_set("-L")
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

@@

Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
\[ref: that CESM run with forced winds\].

Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the eddy-driven jet, which is a crucial component of weather events.
For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
There is some evidence of this effect in the Southern Hemisphere.
Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum.
But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024].
The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location [@ayres2019].
It is also small compared with the direct effect of CO2 increase and Sea Surface Temperature warming [@ayres2019].
On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].

Correctly modelling Antarctic sea undsen sector and the Amundsen-Bellingshausen sector (Fice is not only necessary for process understanding and climate projections to inform adaptation strategies.
Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020].
However, Antarctic sea ice forecasts have been challenging both due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.

{==**S2** Introduce S2==}

{==**Objective.** Evaluate S2.==}

@@

+S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021.
Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
The atmosphere has a N216 horizontal resolution(\~60km in the mid-latitudes) with 85 levels.
The land model uses the same horizontal grid with 4 soil levels.
The ocean component has a 1/4º resolution with 75 vertical levels.
The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.

Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
The main difference between the two are the ocean and ice initial conditions.
+S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022.
This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
Climatology is smoothed with a 11 day running mean.

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 osi = OSI(),
                 era5 = ERA5()
)
which_dataset <- "cdr"
```

There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.

These data are not perfect.
\cite{meier2019} Sensor sensitivity and resolution.
Weather filtering.
Land To account for observational uncertainty we performed the verification on other data sources.
We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
Its sea ice concentration data are estimated taking the maximum value of the NASA Team \[ref\] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
The data has a resolution of 25 by 25km and daily from 1978 onwards.

Results based on two other different datasets are provided as supplementary material:

1.  The NASA Bootstrap [@comiso2023] product is... it is not an independent dataset from NOAA/NSIDC CDRV4.

2.  The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) \[ref\] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.

@@

### Sea ice extent

@@

Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions.
We use two other measures to account for these errors.

We compute Root Mean Square Error of sea ice concentration anomalies.

We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
This is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.

@@

@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
Many S2S systems exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].

Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.

At long lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
The difference between the two is the effect of the data assimilation.

The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to both having the same model formulation.
From June to October, in +S2 circles move away from triangles and towards observations, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and affecting the initial conditions.
The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.

@@

extent_daily |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
  _[dataset == which_dataset] |> 
  ggplot(aes(time, mid)) +
  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
  geom_smooth(aes(group = dataset, color = dataset),
              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
  geom_smooth(data = dif_hindcast,
              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
  scale_x_datetime(NULL, date_breaks = "1 month", 
                   date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_models
```

```{r climatology}
climatologies <- c(datasets,
                   S2 = S2_reanalysis()) |>
  lapply(\(x) climatology(x) |>
           cdo_ymonmean() |> 
           cdo_execute()
  ) |>
  lapply(\(x) ReadNetCDF(x, "aice")) |>
  rbindlist(idcol = "model") 
```

@@

obs_clim <- CDR() |> 
  climatology() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  setnames("time", "time2")

@@

```{r fig-bias}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: 1-month lag +S2 forecast bias compared with NSIDC.

@@

@fig-bias shows the difference in monthly mean sea ice concentrations between +CDR and +S2 reanalysis.
From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1.
In winter, the differences are mostly on the sea ice edge, with slight positive bias in XXX and negative bias around the Indian Ocean sector.

@@

{==intro connecting subsections==}

@@

@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.

@@

@@

errors |> 
  _[measure == "rmse"] |>
  _[obs_dataset ==  which_dataset] |>
  _[, median_ci(value), 
    by = .(lag, version, measure, month(time_forecast), obs_dataset)] |> 
  _[version == "persistence", version := obs_dataset] |> 
  ggplot(aes(lag, mid)) +

@@

  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
              alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(y = "RMSE", 
       x = "Lag") +
  scale_color_models +
  scale_fill_models +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
```

@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a reference?=
=}.
Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
The only exception is the RMSE around February.

@@

```{r rmse_lon}
rmse_lon <- here::here("data/derived/rmse_lon.Rds") |> 
  readRDS() |> 
  _[version != "S1"] |> 
  _[, lag := as.numeric(as.Date(time) - time_forecast)]

rmse_lon[, base := value[version == "persistence"], 
         by = .(time, lon, lag, obs_dataset)]

rmse_lon_mean <- rmse_lon |> 
  _[version != "persistence"] |>
  _[, .(dif = median(value - base, na.rm = TRUE)),
    by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
```

@@

  scale_y_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(fill = NULL, 
       y = "Lag") +
  wide_legend +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
  theme(panel.background = element_blank()) +
  tag_facets(position = list(x = 0.1, y = 1-.02))
```

Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions. 
To analyse the spatial distribution of the model error, we computed ice concentration RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]` degrees thick.
The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.

The skill shown by +S2 at February-March forecasts is evident by a band of negative values across almost all longitudes except around and east of the Antarctic Peninsula, where most of the Summer ice is left. 
This suggests that the positive skill comes from the model correctly forecasting no ice where there is never ice. 
However, there are other regions of skilful forecasts.

Forecasts initialised on January 1^st^ ([Fig. @fig-rmse_lon] panel a) shows RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea. 
These two regions are also relatively well forecasted from other months. 
Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skillful after about 20 days and remain so until December. 
The Weddell sea region is also the region with the maximum erro --arround June irregardless of initialisation date.

@@

geom_line(aes(color = version), linewidth = 1) +
scale_y_continuous("Variance", 
labels = labels_extent_2) +
# scale_color_models +
scale_x_continuous("Lag", breaks = NULL) +
scale_color_models +
facet_grid(month ~ component, 
labeller = labeller(month = labels_month, 
component = c(mean_variance = "Mean variance",
variance_of_mean = "Variance of mean"))) 
```

>>>>

---
title: Evaluation of +S2 sea ice forecast and some other better title
pdf-engine: pdflatex
journal:
  name: tc
author:
  - given_name: Elio 
    surname: Campitelli
    affiliation: "1,2"
    email: elio.campitelli@monash.edu
    corresponding: true
  - given_name: Ariaan 
    surname: Purich
    affiliation: "1,2"
  - given_name: Julie
    surname: Arblaster
    affiliation: "1,2"
  - given_name: Eun-Pa
    surname: Lim
    affiliation: "3"
  - given_name: Matthew
    surname: Wheeler
    affiliation: "3"
  - given_name: Phillip
    surname: Reid
    affiliation: "3"

affiliation:
  - code: 1
    address: School of Earth, Atmosphere and Environment, Monash University, Australia
  - code: 2
    address: Securing Antarctica’s Environmental Future, Monash University, Australia
  - code: 3
    address: Bureau of Meteorology

abstract: |
  The abstract goes here.
  It can also be on _multiple lines_.
bibliography: references.bib
running:
  title: +S2 sea ice forecast
  author: Campitelli et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company. 
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #code: |
  #  use this to add a statement when having only software code available
  #data: |
  #  use this to add a statement when having only data sets available
  codedata: |
    use this to add a statement when having data sets and software code available
  sample: |
    use this section when having geoscientific samples available
videosupplement: |
  use this section when having video supplements available
authorcontribution: |
  Contributions
disclaimer: |
  We like Copernicus.
acknowledgements: |
  Acknowledgements will gohere
appendix: |
  \section{Figures and tables in appendices}
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  \subsection{Option 1}
  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
  They will be correctly named automatically.
  \subsection{Option 2}
  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
  `\appendixfigures` needs to be added in front of appendix figures
  `\appendixtables` needs to be added in front of appendix tables

  Please add `\clearpage` between each table and/or figure.
format: copernicus-pdf

execute: 
  echo: false
  warning: false
  message: false
  cache: refresh
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  # - ../../_extensions/mloubout/critic-markup/critic-markup.lua
  - ../../_extensions/ute/search-replace/search-replace.lua
search-replace:
  +S2: ACCESS-S2
  +S1: ACCESS-S1
  +CDR: NOAA/NSIDC CDRV4
---

```{r setup, include=FALSE, cache = FALSE}
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)
library(tagger)

cdo_options_set("-L")
```

```{r helpers, cache = FALSE}
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

@@

Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change [@espinosa2024].

Antarctic sea ice is in a remote location but the potential impacts of its variability extends far into the lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes.
It's likely that a reduction in sea ice extent would reduce the meridional temperature gradient which in turn would reduce the strength and location of the eddy-driven jet (and vice versa for an increase in sea ice), which is a crucial component of weather events.
In the northern hemisphere, it has been hypothesised that the reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
There is some evidence of this effect in the Southern Hemisphere.
Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum.
But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024].
The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location and relatively small compared with the direct effect of CO2 increase and Sea Surface Temperature (SST) warming [@ayres2019].
On climate scales, it is thought that the temperature signal of a reduction in Antarctic sea ice mirrors a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].

Correctly modelling Antarctic sea  is not only necessary for process understanding and climate projections to inform adaptation strategies.
Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.

In this work we evaluate sea ice forecasts produced by +S2, the Australian Bureau of Meteorology subseasonal-to-seasonal prediction system. 
This evaluation will inform future work with the model as a reseach tool and explore the potential of using it as a sea ice forecasting tool.

@@

+S2 [@wedd2022] is the Australian Bureau of Meteorology current subseasonal-to-seasonal prediction system which replaced +S1 [@hudson2017] in October 2021.
Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015] model components.
The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 levels.
The land model uses the same horizontal grid with 4 soil levels.
The ocean component has a 1/4º resolution with 75 vertical levels.
The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.

Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
The main difference between the two are the ocean and ice initial conditions.
+S1 initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022.
This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

The +S2 hindcast used in this study runs for the period 1981--2018.
Each forecast consists of 9 ensemble members, 3 of which run for 279 days and 6 for 42 days.
Anomalies are be taken with respect to the 1981--2011 climatology computed from the daily +S2 reanalysis and smoothed with a 11 day running mean.

The +S1 hindcast runs for the period 1990--2012, with 7-member forecasts integrated for 217 days. 
No reanalysis is available so we are not able to compute a climatology for this forecast, therefore we don't show metrics based on anomalies.

@@

```{r datasets}
datasets <- list(cdr = CDR(),
                 osi = OSI())
which_dataset <- "cdr"
```

For verification we use satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
Different datasets derived using different algorithms provide different estimates, each with its own biases and uncertainties. 
@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million km^2^. 
As will be shown below, this spread is minimal compared with the errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
It takes the maximum value of the NASA Team [@cavalieri1984] and NASA Bootstrap [@comiso2023] sea ice concentration products to reduce their low concentration bias [@meier2014; @meier2021].
Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
The data has a resolution of 25 by 25km and daily from 1978 onwards.

The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [@OSISAF] is another satellite-derived sea ice concentration product. 
It is based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
Figures prepared with this dataset are provided in the supplementary material.

@@

@@

Pan-Antarctic Sea Ice Extent serves as a rough global measure of the amount of sea ice but it doesn't take into account the spatial distribution. 
A model could have relatively accurate extent of ice but with different distributions. To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].

We compute RMSE as the square root of the average squared differences between forecasted and observed sea ice concentration anomalies. 
We compute a pan-Antarctic RMSE by averaging over the whole domain, and also a zonally-varying RMSE by averaging over 24 15° longitude slices around Antarctica.

The IIEE is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.

All error measures were computed on the +CDR domain grid and projection to which model output was bilinearly interpolated.

@@

@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April. 
This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].

+S1 has a smaller low bias than +S2, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1 and not in +S2.

At large lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
The difference between the two is the effect of the data assimilation.

The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to them having the same model formulation.
From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and improving the initial conditions.
The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
In +S1, on the other hand, circles are very close to observations year-round.

@@

extent_daily |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
  _[dataset == which_dataset] |> 
  ggplot(aes(time, mid)) +
  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
  geom_smooth(aes(group = dataset, color = dataset),
              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
  geom_smooth(data = dif_hindcast,
              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
  scale_x_datetime(NULL, date_breaks = "1 month", 
                   date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = \(x) paste0(labels_extent(x), "/day")) +
  scale_color_models
```

@@

obs_clim <- CDR() |> 
  climatology() |> 
  cdo_ymonmean() |>
  cdo_execute() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  setnames("time", "time2") |> 
  _[, time2 := update(time2, day = 1)]

@@

```{r fig-bias}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.

@@

@fig-bias shows the difference of monthly mean sea ice concentrations between +CDR and +S2 hindcasts at the shortest lag. 
From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
In winter, the differences are mostly on the sea ice edge, with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig. @fig-hindcast-extent]).

@@

@@

@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
The anomalies in this case were computed with respect of the climatology of each lag, which works as a bias-correction method. 
Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.

@@

```{r clim_std}
clim_std <- datasets |> 
  lapply(\(x) x |> 
           anomalies() |> 
           cdo_sqr() |> 
           cdo_fldmean() |> 
           cdo_sqrt() |> 
           cdo_execute() |> 
           ReadNetCDF(c(std = "aice")) |> 
           _[, let(lat = NULL, lon = NULL)]
  ) |> 
  rbindlist(idcol = "obs_dataset") |> 
  _[, time := as.Date(time)] |> 
  _[std == 0, std := NA] |> 
  _[std < 0.05, std := NA]

@@

errors |> 
  _[measure == "rmse"] |>
  _[, time2 := update(time, year = 2000)] |> 
  # merge(clim_std) |>
  _[obs_dataset ==  which_dataset] |>
  _[, median_ci(value),
    by = .(lag, version, measure, month(time_forecast), obs_dataset)] |>
  _[version == "persistence", version := obs_dataset] |> 
  ggplot(aes(lag, mid)) +

@@

  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
              alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +

  # geom_line(aes(y = clim_std)) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(y = "RMSE", 
       x = "Lag") +
  scale_color_models +
  scale_fill_models +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
```

@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, after when it is statistically useful.
Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lag.

@@

```{r rmse_lon_mean}
rmse_lon_mean <- here::here("data/derived/rmse_lon.Rds") |> 
  readRDS() |> 
  _[version != "S1"] |> 
  _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
  _[, base := value[version == "persistence"], 
    by = .(time, lon, lag, obs_dataset)] |> 
  _[version != "persistence"] |>
  _[, .(dif = median(value - base, na.rm = TRUE)),
    by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
```

@@

  scale_y_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(fill = NULL, 
       y = "Lag") +
  wide_legend +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
  theme(panel.background = element_blank(), 
        tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +
  tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
```

Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions. 
To analyse the spatial distribution of the model error, we computed sea ice concentration anomalies RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.

The skill shown by +S2 at February-March forecasts is evidenced by a band of negative values across almost all longitudes. 
The only region where February-March sea ice is not well forecasted is east of the Antarctic Peninsula, which is the only region with consistent summer sea ice.
This suggests that the positive skill around the summer minimum comes from the model correctly forecasting no ice where no ice is ever found. 
However, there are other regions of skilful forecasts.

Forecasts initialised on January 1^st^ ([Fig. @fig-rmse_lon] panel a) show RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea. 
These two regions are also relatively well forecasted from other months. 
Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skilful after about 20 days and remain so until December. 
The Weddell Sea is also the region with the maximum error --around June regardless of initialisation date.

@@

  geom_line(aes(color = version), linewidth = 1) +
  scale_y_continuous("Variance", 
                     labels = labels_extent_2) +
  # scale_color_models +
  scale_x_continuous("Lag", breaks = NULL) +
  scale_color_models +
  ggh4x::facet_nested_wrap(month ~ component, ncol = 6, 
                           remove_labels = "all",
             labeller = labeller(month = labels_month, 
                                 component = c(mean_variance = "Mean variance",
                                               variance_of_mean = "Variance of mean"))) 
```
